### 2016-11-09 datamining/summery.tex
Finished with freq pattern mining

\documentclass[10pt,landscape]{extarticle}
\usepackage{multicol}
\usepackage{calc}
\usepackage{ifthen}
\usepackage[landscape]{geometry}
\usepackage{amsmath,amsthm,amsfonts,amssymb}
\usepackage{color,graphicx,overpic}
\usepackage{hyperref}
\usepackage{graphicx}
\newcommand{\indep}{\rotatebox[origin=c]{90}{$\vDash$}}

@@

\begin{center}
     \Large{\underline{Datamining}} \\
\end{center}
\section{Classification trees}
\subsection{Building classification trees}
The objective of tree construction is to finally obtain nodes that are pure,
in the sense that they contain cases of a single class only.
\subsubsection{Impurity and quality of a split}
Impurity is a function:
\[i(t)=\phi(p_1,p_2,...,p_J)\]
where $p_j(j=1,...,J)$ is the relative frequency of the $J$ different classes.
$\phi$ is the impurity function with the following constraints:
\begin{enumerate}
  \item An impurity measure of a node should be at maximum when the observations
    are distributed evenly over all classes in that node.
  \item It should be at a minimum when all observations belong to a single class.
  \item $\phi$ is a symmetric function of $p_1,...,p_J$.
    a symmetric function of n variables is one whose value at any n-tuple of
    arguments is the same as its value at any permutation of that n-tuple.
\end{enumerate}
The quality of a split is defined as follows:
\[\Delta i(s,t)=i(t)-\phi(l)i(l)-\phi(r)i(r)\]
Where $\phi$ is a function that determines the proportion of cases for a node
and $l$ and $r$ are the left and right nodes of $t$.
\paragraph{Resubstitution error}
It measures what fraction of the class in a node is classified incorrectly if
we assign every case to the majority class in that node.
\paragraph{Gini} $i(t)=\sum_j p(j|t)(1-p(j|t))$
\paragraph{entropy}$i(t)=\sum_j p(j|t)\log p(j|t)$

@@

>>>>

\documentclass[9pt,landscape]{extarticle}
\usepackage{multicol}
\usepackage{calc}
\usepackage{ifthen}
\usepackage[landscape]{geometry}
\usepackage{amsmath,amsthm,amsfonts,amssymb}
\usepackage{color,graphicx,overpic}
\usepackage{hyperref}
\usepackage{graphicx}
\newcommand{\indep}{\rotatebox[origin=c]{90}{$\vDash$}}
\usepackage{amssymb}
\usepackage{algorithmicx}
\usepackage{algpseudocode}

@@

\begin{center}
     \Large{\underline{Datamining}} \\
\end{center}
\section{Classification trees}
\subsection{Building classification trees}
The objective of tree construction is to finally obtain nodes that are pure,
in the sense that they contain cases of a single class only.
\subsubsection{Impurity and quality of a split}
Impurity is a function:
$i(t)=\phi(p_1,p_2,...,p_J)$
where $p_j(j=1,...,J)$ is the relative frequency of the $J$ different classes.
$\phi$ is the impurity function with the following constraints:
An impurity measure of a node should be at maximum when the observations
are distributed evenly over all classes in that node,
it should be at a minimum when all observations belong to a single class,
$\phi$ is a symmetric function of $p_1,...,p_J$.
A symmetric function of n variables is one whose value at any n-tuple of
arguments is the same as its value at any permutation of that n-tuple.
The quality of a split is defined as follows:
\[\Delta i(s,t)=i(t)-\phi(l)i(l)-\phi(r)i(r)\]
Where $\phi$ is a function that determines the proportion of cases for a node
and $l$ and $r$ are the left and right nodes of $t$.
\paragraph{Resubstitution error}
It measures what fraction of the class in a node is classified incorrectly if
we assign every case to the majority class in that node.
\paragraph{Gini} $i(t)=\sum_j p(j|t)(1-p(j|t))$
\paragraph{entropy}$i(t)=\sum_j p(j|t)\log p(j|t)$

@@

\subsection{Association rules: The binary case}
The traditional setting for association rules consists of a set of transactions
in which each transaction is a set of items.
Translated to a relational setting,
we have a table db with schema $R = {I_1 , \dots, I_n }$,
in which each $I_i$ is a binary attribute.
The attributes correspond with the items,
the rows in the table with the transactions
; a row has value 1 for an attribute if and only if the transaction contains
that item.
For $X,Y \subseteq R$, with $X\cap Y=\emptyset$ let $s(X)$ denote the support of
$X$, the number of tuples that have value 1 for all items in $X$. for an
association rule $X \to Y$ define the support is $s(X\cup Y)$, the confidence is
$s(X\cup Y)/s(X)$
The problem is to find all association rules that match or exceed the user
defined lower thresholds for confidence, \emph{minconf}, and support
\emph{minsup}.
In a major abuse of notation,
you could say that the support of X (when expressed as a fraction) corresonds
to $P (X)$ in probability terms,
and the confidence of $X \to Y$ with the conditional probability $P (Y | X)$.
There are two tesholds we have to satisfy:
find all sets $Z$ whoese support exceeds the minimal threshold. These sets are
called frequent. Then test for all non-empty subsets $X$ of frequent sets $Z$
wheter the rul $X \to Z \setminus X(=Y)$ holds with sufficient confidence.
\subsubsection{A priori alogirhtm}
Finding all frequents sets is problematic since all sets of $R$ are
$|\mathcal{P}(R)|=2^{|R|}$. But since $Z$ can only be frequent if all its
(non-empty) subsets are frequent we can do it.
This is known as the A-Priori property. In other words, we can search level wise
for the frequent sets.
The level is the number of items in the set. Denote by $C(i)$ the sets of $i$
items that are potentially frequent (the candidate sets) and by $F (i)$
the frequent sets of $i$ items.
Start with level one all entries of $R$ (in random order),
filter on the level $C(i)$ each entry
that doesn't have enough support, these are the frequent items $F(i)$.
Then $i$ is increased, and we start creating the
next level of candidates $C(i)$ based on $F(i-1)$ (which we just made).
Then they do this:
For each $X \in F (i − 1)$ do
For each $Y \in F (i − 1)$ that shares the first $i − 2$ items with X do
$C(i) := C(i) \cup {X \cup Y }$
If we assume the db is sparse the complexity becomes $O(|R|^{k+1})$ where $k$
is the maximal size of frequent items sets that is $k << |R|$.
To generate association rules:
\emph{flatmap} each freq set $X$ \emph{with} each non empty $Y \subset X$ filtering
$s(X)/s(X\setminus Y)\geq minconf$ as $X \setminus Y \to Y $
\subsection{Drowining in assoc rules}
\subsubsection{Postprocessing the result set}
We can order the resultss we can use conidence and support: they define a
partial order on the set of rules. Given rules $r_1$ and $r_2, r_1A$ iff 
$s(r_1) \leq s(r_2) \wedge conf(r_1) < con (r_2)$ or
$s(r_1) < s(r_2) \wedge conf(r_1) \leq conf(r_2)$
A rule is interesting if it gives usefull infromation. Lots of measures have
been defined as an attempt to do that: lift, interest, conviction, colective
strength, gain, gini, entropy, $\chi^2$ This is lift:
$lift(A\to C)=\frac{P(C|A)}{P(C)}=\frac{P(A,C)}{P(A)P(C)}$
The lift should be bigger than one for a rle to be itneresting.
\subsubsection{Generating less rules}
Closed frequent itemsets are itemsets that completely characterise their
associated set of transactions.
That is, a frequent itemset $I$ is closed if $I$ contains all items that occur in
all transactions in the support of $I$.
For an itemset $I$, denote by $\sigma(I)$ the set of tuples in which all items
in $I$ are bought, that is $\sigma(I)=\{t \in db | \forall i \in I, i \in t\}$
For a set of transaction $T$ let $f(T)$ denote the set of items that are bought
in all transaction in $T$, that is $f(T)=\{i \in R | \forall t \in T, i \in t\}$
The closure can be obtained by $c(I)=f(\sigma(I))$.
To give a procedural description of applying the closure operator to an itemset
$I$:
first get the transactions in which all items in $I$ are bought,
and then see whether there are any more items that are common to all these
transactions.
If so, add them to $I$ and return the result.
Clearly, $I \subseteq c(I)$ and $I$ has the same support as $c(I)$.
An itemset $I$ is closed if and only if $c(I) = I$.
An itemset $I$ is a closed frequent itemset iff it is both frequent and closed.
The A-Close algorithm for finding al frequent closed itemsets consists of 2
phases: One, discover all freq closed itemsets in db. Derive all frequent
itemsets from the freqeunt closed itemsets found in phase 1.
In phase 1 the algorithm determines a set of so-called generators that will
produce all frequent closed itemsets by application of the closure operator $c$.
An itemset $I$ is a generator of a closed itemset $J$ if it is one of the
smallest itemsets with $c(I) = J$.
A-Close performs a levelwise search like Apriori:
$G_{i+1}$ (the set of generators at level $i + 1$) is constructed using $G_i$.
Using their support, and the support of their $i$-subsets in $G_i$,
infrequent itemsets and itemsets that have the same support as one of their
subsets are deleted from $G_{i+1}$. The rational of the secon pruning rule is
that if an itemset has a subset with the same suppport, then this subset also
has the same closure.
Now that the generators have been determined, we compute their closures,
where $c(I)=\cap t \in db: I \subseteq t$. Different generators may produce the
same closure, duplicates are removed.
In phase 2 we determine all frequent itemsets and their support.
We use the following properties:
1. All maximal frequent itemsets are closed.
2. Every frequent itemset is a subset of a maximal frequent itemset.
3. The support of an itemset equals the support of the smallest closed itemset
in which it is contained (its closure).
Our strategy is therefore to select the maximal itemsets,
generate all their sub- sets,
and determine their support using the third property.
